---
title: | 
  | Portable Proteomics Pipeline (P3) 
  | MSGF Benchmark
author: "Daniel Kristiyanto (daniel.kristiyanto@pnnl.gov)"
date: \today
fontsize: 11pt
output: pdf_document
# bibliography: ../../media/ref.bib
nocite: |
  @domon2006mass, @deutsch2010mass, @MSnba83
csl: ../../media/biomed-central.csl
header-includes:
  \usepackage{fancyhdr}
  \usepackage{graphicx}
  \usepackage{hyperref}
  \pagestyle{fancy}
  \setlength\headheight{28pt}
  \rhead{\includegraphics[width = .2\textwidth]{../../media/pnnllogo.png}}
  \fancyfoot[LE,RO]{kristiyanto/p3:benchmark}
---
\rule{\textwidth}{1pt}
```{r, warning=FALSE, echo=FALSE, include=FALSE, cache=T}
library(MSnbase)
library(mzID)
library(stringr)
suppressMessages(library(BiocParallel))

```

```{r, warning=FALSE, echo=FALSE, include=FALSE, cache=T}
setwd("~/Dropbox/_SPACE/Benchmark/")
file_mzid        <- list.files(path = ".", pattern ="mzid", all.files = F, 
                      full.names = F, recursive = F, ignore.case = T, include.dirs = F)
file_mzml        <- list.files(path = ".", pattern ="mzML$|mzXML$", all.files = F, 
                      full.names = F, recursive = F, ignore.case = T, include.dirs = F)
file_fasta        <- list.files(path = ".", pattern ="fasta", all.files = F, 
                      full.names = F, recursive = F, ignore.case = T, include.dirs = F)

mzid_pnnl <- read.table(file="TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_msgfdb_fht.txt", sep="\t", header=T)
mzids_no_switch   <- mzID("TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_no_switch.mzid")
mzid_p3           <- mzID("TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.mzid")
tsv_ns <- read.table(file ="TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_no_switch.tsv", sep="\t",header = T)
tsv <- read.table(file ="TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.tsv", sep="\t",header = T)

msexp.raw         <- readMSData(file_mzml)
msexp_p3          <- addIdentificationData(msexp.raw, id = mzid_p3,verbose=F)
msexp_no_switch   <- addIdentificationData(msexp.raw, id = mzids_no_switch,verbose=F)

p3_table <- fData(msexp_p3)
ns_table <- fData(msexp_no_switch)

pnnl <- mzid_pnnl[, c("Scan", "MSGFDB_SpecEValue", "Peptide")]
pnnl <- pnnl[!duplicated(pnnl$Scan),]
p3_flt <- flatten(mzid_p3)
bigmzid <- merge(pnnl, p3_table[, c("scan number(s)", "ms-gf:specevalue", "pepseq")], by.x="Scan", by.y = "scan number(s)", all.x = T, all.y = T)
pnnl$Peptide <- substr(as.character(pnnl$Peptide),1,nchar(as.character(pnnl$Peptide))-2)
pnnl$Peptide <- substr(as.character(pnnl$Peptide),3,nchar(as.character(pnnl$Peptide)))

```

# 1. Files:

```{r, echo=F}
show(file_mzml)
show(file_fasta)
```

# 2. Identification:

`MSGFPlus.jar` was downloaded from \url{https://omics.pnl.gov/software/ms-gf}. Identification was run multiple times (within or without Docker), and resulted the same results. 

Identification with different switches resulted different results.

```
# Command Line (Without Switch)
java -Xmx8000M -jar P3/MSGFPlus.jar -s \
TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.mzML -o \
TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_no_switch.mzid \
-d ID_003632_9011437E.fasta  

# Command Line (2)
java -Xmx8000M -jar P3/MSGFPlus.jar -s \
TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.mzML -o \
TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11.mzid \
-d ID_003632_9011437E.fasta  \
-t 10ppm -m 0 -inst 1 -e 1 -ti -1,2 -ntt 2 -tda 1 -minLength 6 \
-maxLength 50 -minCharge 2 -maxCharge 5 -n 1 -thread 7 \
-mod MSGFDB_Mods.txt -minNumPeaks 5 -addFeatures 1
```
```{r, echo=F}
# Identification Result
show(file_mzid)
```

```{r, echo = F, include=F}
evalue = 1e-10
```

#3. Result from PNNL
```{r, echo=F}
# Nrow
nrow(pnnl)
head(pnnl,2)
tail(pnnl,2)
```


#4. Result from MSGF+ without any switches (sorted by SpecEValue)
```{r, echo=F}
tmp           <- ns_table
tmp           <- tmp[order(tmp$'ms-gf:specevalue'),]
show(head(tmp[,c("scan number(s)", "ms-gf:denovoscore", "ms-gf:specevalue",  "pepseq")])[1:3,])
## Looking at Scan Number 11612 and 14441
tmp[(tmp$`scan number(s)`=="11612" | tmp$`scan number(s)`=="14441")  &! is.na(tmp$`scan number(s)`),c("scan number(s)", "ms-gf:denovoscore", "ms-gf:specevalue", "pepseq")]
```


Reading the MZID file manually
```
java -Xmx2000M -XX:+UseConcMarkSweepGC -cp \ 
../P3/MSGFPlus.jar edu.ucsd.msjava.ui.MzIDToTsv -i \
TCGA_13-1489_42-2590_36-2529_117C_W_PNNL_B2S5_f11_no_switch.mzid \
-showQValue 1 -showDecoy 1 -unroll 1
```

```{r, echo=F}
head(tsv_ns)[1:3,c("ScanNum", "Peptide", "Protein", "SpecEValue")]
```


# 5. Results with switches
```{r, echo=F}
tmp           <- p3_table
tmp           <- tmp[order(tmp$'ms-gf:specevalue'),]
show(head(tmp[,c("scan number(s)", "ms-gf:denovoscore", "ms-gf:specevalue", "pepseq")])[1:3,])
## Looking at Scan Number 11612 and 14441
tmp[(tmp$`scan number(s)`=="11612" | tmp$`scan number(s)`=="14441") &! is.na(tmp$`scan number(s)`),c("scan number(s)", "ms-gf:denovoscore", "ms-gf:specevalue", "accession", "pepseq")]

```


Reading the MZID file manually
```{r, echo=F}
head(tsv)[1:3,c("ScanNum", "Peptide", "SpecEValue")]
```

# 6. Filtering (Docker --> PNNL)

```{r include=F}
evalue = 1e-10
```

# a Spec-Evalue `r evalue`
### From PNNL
```{r, echo =F}
pnnl_tmp <- pnnl[pnnl$MSGFDB_SpecEValue<evalue,]
print("==Head==")
head(pnnl_tmp,2)
print("==Tail==")
tail(pnnl_tmp,2)
```
There are ```r nrow(pnnl_tmp)``` rows remained.

### From Docker 
```{r echo=F}
tmp <- (p3_table[p3_table$'ms-gf:specevalue'<evalue & !is.na(p3_table$'ms-gf:specevalue'<evalue),c("scan number(s)", "ms-gf:specevalue", "pepseq")])
tmp <- tmp[order(tmp$`ms-gf:specevalue`),]
print("==Head==")
head(tmp,2)
print("==Tail==")
tail(tmp,2)

bt_tmp <- merge(x =  tmp, y=pnnl, by.y = "Scan", by.x  =  "scan number(s)", all.x = T)
bt_tmp <- bt_tmp[order(bt_tmp$`ms-gf:specevalue`, decreasing = F),]
tmp_nona <- bt_tmp[is.na(bt_tmp$MSGFDB_SpecEValue),]
tmp_nona <- tmp_nona[order(tmp_nona$'ms-gf:specevalue', decreasing = F),]
bt_tmp <- bt_tmp[!is.na(bt_tmp$Peptide),]
```
There are ```r nrow(tmp)``` rows remained.

### Differences
Data from Docker is matched with from PNNL using scan number as ID. `r names(bt_tmp)[2:3]` are values from Docker. `r names(bt_tmp)[4:5]` are the corresponding values from PNNL. There are `r nrow(bt_tmp)` of rows in total with  `r nrow(bt_tmp[((bt_tmp$'ms-gf:specevalue'==bt_tmp$MSGFDB_SpecEValue) & (bt_tmp$pepseq==bt_tmp$Peptide)),])` exact match.

Aside from the `r nrow(tmp_nona)` unmatched values, `r sum(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue)` have different spec-evalue, `r sum(bt_tmp$pepseq!=bt_tmp$Peptide)` of different peptide identification.

```{r echo=F}
print("==Head==")
head(bt_tmp,2)
print("==Tail==")
tail(bt_tmp,2)

print("==Unmatched==")
head(tmp_nona,2)
print("==Different Spec Values==")
head(bt_tmp[(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue),],2)
print("==Different Peptides identification==")
head(bt_tmp[(bt_tmp$pepseq!=bt_tmp$Peptide),],2)

```

```{r include=F}
evalue = 1e-15
```

# b. Spec-Evalue `r evalue`
### From PNNL
```{r, echo =F}
pnnl_tmp <- pnnl[pnnl$MSGFDB_SpecEValue<evalue,]
print("==Head==")
head(pnnl_tmp,2)
print("==Tail==")
tail(pnnl_tmp,2)
```
There are ```r nrow(pnnl_tmp)``` rows remained.

### From Docker 
```{r echo=F}
tmp <- (p3_table[p3_table$'ms-gf:specevalue'<evalue & !is.na(p3_table$'ms-gf:specevalue'<evalue),c("scan number(s)", "ms-gf:specevalue", "pepseq")])
tmp <- tmp[order(tmp$`ms-gf:specevalue`),]
print("==Head==")
head(tmp,2)
print("==Tail==")
tail(tmp,2)

bt_tmp <- merge(x =  tmp, y=pnnl, by.y = "Scan", by.x  =  "scan number(s)", all.x = T)
bt_tmp <- bt_tmp[order(bt_tmp$`ms-gf:specevalue`, decreasing = F),]
tmp_nona <- bt_tmp[is.na(bt_tmp$MSGFDB_SpecEValue),]
tmp_nona <- tmp_nona[order(tmp_nona$'ms-gf:specevalue', decreasing = F),]
bt_tmp <- bt_tmp[!is.na(bt_tmp$Peptide),]
```
There are ```r nrow(tmp)``` rows remained.

### Differences
Data from Docker is matched with from PNNL using scan number as ID. `r names(bt_tmp)[2:3]` are values from Docker. `r names(bt_tmp)[4:5]` are the corresponding values from PNNL. There are `r nrow(bt_tmp)` of rows in total with  `r nrow(bt_tmp[((bt_tmp$'ms-gf:specevalue'==bt_tmp$MSGFDB_SpecEValue) & (bt_tmp$pepseq==bt_tmp$Peptide)),])` exact match.

Aside from the `r nrow(tmp_nona)` unmatched values, `r sum(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue)` have different spec-evalue, `r sum(bt_tmp$pepseq!=bt_tmp$Peptide)` of different peptide identification.

```{r echo=F}
print("==Head==")
head(bt_tmp,2)
print("==Tail==")
tail(bt_tmp,2)

print("==Unmatched==")
head(tmp_nona,2)
print("==Different Spec Values==")
head(bt_tmp[(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue),],2)
print("==Different Peptides identification==")
head(bt_tmp[(bt_tmp$pepseq!=bt_tmp$Peptide),],2)

```


```{r include=F}
evalue = 1e-20
```

# c. Spec-Evalue `r evalue`
### From PNNL
```{r, echo =F}
pnnl_tmp <- pnnl[pnnl$MSGFDB_SpecEValue<evalue,]
print("==Head==")
head(pnnl_tmp,2)
print("==Tail==")
tail(pnnl_tmp,2)
```
There are ```r nrow(pnnl_tmp)``` rows remained.

### From Docker 
```{r echo=F}
tmp <- (p3_table[p3_table$'ms-gf:specevalue'<evalue & !is.na(p3_table$'ms-gf:specevalue'<evalue),c("scan number(s)", "ms-gf:specevalue", "pepseq")])
tmp <- tmp[order(tmp$`ms-gf:specevalue`),]
print("==Head==")
head(tmp,2)
print("==Tail==")
tail(tmp,2)

bt_tmp <- merge(x =  tmp, y=pnnl, by.y = "Scan", by.x  =  "scan number(s)", all.x = T)
bt_tmp <- bt_tmp[order(bt_tmp$`ms-gf:specevalue`, decreasing = F),]
tmp_nona <- bt_tmp[is.na(bt_tmp$MSGFDB_SpecEValue),]
tmp_nona <- tmp_nona[order(tmp_nona$'ms-gf:specevalue', decreasing = F),]
bt_tmp <- bt_tmp[!is.na(bt_tmp$Peptide),]
```
There are ```r nrow(tmp)``` rows remained.

### Differences
Data from Docker is matched with from PNNL using scan number as ID. `r names(bt_tmp)[2:3]` are values from Docker. `r names(bt_tmp)[4:5]` are the corresponding values from PNNL. There are `r nrow(bt_tmp)` of rows in total with  `r nrow(bt_tmp[((bt_tmp$'ms-gf:specevalue'==bt_tmp$MSGFDB_SpecEValue) & (bt_tmp$pepseq==bt_tmp$Peptide)),])` exact match.

Aside from the `r nrow(tmp_nona)` unmatched values, `r sum(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue)` have different spec-evalue, `r sum(bt_tmp$pepseq!=bt_tmp$Peptide)` of different peptide identification.

```{r echo=F}
print("==Head==")
head(bt_tmp,2)
print("==Tail==")
tail(bt_tmp,2)

print("==Unmatched==")
head(tmp_nona,2)
print("==Different Spec Values==")
head(bt_tmp[(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue),],2)
print("==Different Peptides identification==")
head(bt_tmp[(bt_tmp$pepseq!=bt_tmp$Peptide),],2)

```

```{r include=F}
evalue = 1e-15
```


# 7. Filtering (PNNL -> Docker)


```{r include=F}
evalue = 1e-10
```

# a. Spec-Evalue `r evalue`
### From PNNL
```{r, echo =F}
pnnl_tmp <- pnnl[pnnl$MSGFDB_SpecEValue<evalue,]
print("==Head==")
head(pnnl_tmp,2)
print("==Tail==")
tail(pnnl_tmp,2)
```
There are ```r nrow(pnnl_tmp)``` rows remained.

### From Docker 
```{r echo=F}
tmp <- (p3_table[p3_table$'ms-gf:specevalue'<evalue & !is.na(p3_table$'ms-gf:specevalue'<evalue),c("scan number(s)", "ms-gf:specevalue", "pepseq")])
tmp <- tmp[order(tmp$`ms-gf:specevalue`),]
print("==Head==")
head(tmp,2)
print("==Tail==")
tail(tmp,2)

bt_tmp <- merge.data.frame(y =  p3_table[,c("scan number(s)", "ms-gf:specevalue", "pepseq")], x=pnnl_tmp, by.x = "Scan", by.y  =  "scan number(s)", all.x = T, all.y = F)
bt_tmp <- bt_tmp[order(bt_tmp$`ms-gf:specevalue`, decreasing = F),]
tmp_nona <- bt_tmp[is.na(bt_tmp$'ms-gf:specevalue'),]
tmp_nona <- tmp_nona[order(tmp_nona$MSGFDB_SpecEValue, decreasing = F),]
bt_tmp <- bt_tmp[!is.na(bt_tmp$'pepseq'),]
```
There are ```r nrow(tmp)``` rows remained.

### Differences
Data from Docker is matched with from PNNL using scan number as ID. `r names(bt_tmp)[2:3]` are values from PNNL. `r names(bt_tmp)[4:5]` are the corresponding values from Docker Container. There are `r nrow(bt_tmp)` of rows in total with   `r nrow(bt_tmp[((bt_tmp$'ms-gf:specevalue'==bt_tmp$MSGFDB_SpecEValue) & (bt_tmp$pepseq==bt_tmp$Peptide)),])` exact match.

Aside from the `r nrow(tmp_nona)` unmatched values, `r sum(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue)` have different spec-evalue, `r sum(bt_tmp$pepseq!=bt_tmp$Peptide)` of different peptide identification.

```{r echo=F}
print("==Head==")
head(bt_tmp,2)
print("==Unmatched==")
head(tmp_nona,3)
print("==Tail==")
tail(bt_tmp,2)


print("==Unmatched==")
head(tmp_nona,2)
print("==Different Spec Values==")
head(bt_tmp[(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue),],2)
print("==Different Peptides identification==")
head(bt_tmp[(bt_tmp$pepseq!=bt_tmp$Peptide),],2)

```

```{r include=F}
evalue = 1e-15
```

# b. Spec-Evalue `r evalue`

### From PNNL
```{r, echo =F}
pnnl_tmp <- pnnl[pnnl$MSGFDB_SpecEValue<evalue,]
head(pnnl_tmp,2)
tail(pnnl_tmp,2)
```
There are ```r nrow(pnnl_tmp)``` rows remained.

### From Docker 
```{r echo=F}
tmp <- (p3_table[p3_table$'ms-gf:specevalue'<evalue & !is.na(p3_table$'ms-gf:specevalue'<evalue),c("scan number(s)", "ms-gf:specevalue", "pepseq")])
tmp <- tmp[order(tmp$`ms-gf:specevalue`),]
print("==Head==")
head(tmp,2)
print("==Tail==")
tail(tmp,2)

bt_tmp <- merge.data.frame(y =  p3_table[,c("scan number(s)", "ms-gf:specevalue", "pepseq")], x=pnnl_tmp, by.x = "Scan", by.y  =  "scan number(s)", all.x = T, all.y = F)
bt_tmp <- bt_tmp[order(bt_tmp$`ms-gf:specevalue`, decreasing = F),]
tmp_nona <- bt_tmp[is.na(bt_tmp$'ms-gf:specevalue'),]
tmp_nona <- tmp_nona[order(tmp_nona$MSGFDB_SpecEValue, decreasing = F),]
bt_tmp <- bt_tmp[!is.na(bt_tmp$'pepseq'),]
```
There are ```r nrow(tmp)``` rows remained.

### Differences
Data from Docker is matched with from PNNL using scan number as ID. `r names(bt_tmp)[2:3]` are values from PNNL. `r names(bt_tmp)[4:5]` are the corresponding values from Docker Container. There are `r nrow(bt_tmp)` of rows in total with   `r nrow(bt_tmp[((bt_tmp$'ms-gf:specevalue'==bt_tmp$MSGFDB_SpecEValue) & (bt_tmp$pepseq==bt_tmp$Peptide)),])` exact match.

Aside from the `r nrow(tmp_nona)` unmatched values, `r sum(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue)` have different spec-evalue, `r sum(bt_tmp$pepseq!=bt_tmp$Peptide)` of different peptide identification.

```{r echo=F}
print("==Head==")
head(bt_tmp,2)
print("==Unmatched==")
head(tmp_nona,3)
print("==Tail==")
tail(bt_tmp,2)


print("==Unmatched==")
head(tmp_nona,2)
print("==Different Spec Values==")
head(bt_tmp[(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue),],2)
print("==Different Peptides identification==")
head(bt_tmp[(bt_tmp$pepseq!=bt_tmp$Peptide),],2)

```



```{r include=F}
evalue = 1e-20
```

# c. Spec-Evalue `r evalue`

### From PNNL
```{r, echo =F}
pnnl_tmp <- pnnl[pnnl$MSGFDB_SpecEValue<evalue,]
head(pnnl_tmp,2)
tail(pnnl_tmp,2)
```
There are ```r nrow(pnnl_tmp)``` rows remained.

### From Docker 
```{r echo=F}
tmp <- (p3_table[p3_table$'ms-gf:specevalue'<evalue & !is.na(p3_table$'ms-gf:specevalue'<evalue),c("scan number(s)", "ms-gf:specevalue", "pepseq")])
tmp <- tmp[order(tmp$`ms-gf:specevalue`),]
print("==Head==")
head(tmp,2)
print("==Tail==")
tail(tmp,2)

bt_tmp <- merge.data.frame(y =  p3_table[,c("scan number(s)", "ms-gf:specevalue", "pepseq")], x=pnnl_tmp, by.x = "Scan", by.y  =  "scan number(s)", all.x = T, all.y = F)
bt_tmp <- bt_tmp[order(bt_tmp$`ms-gf:specevalue`, decreasing = F),]
tmp_nona <- bt_tmp[is.na(bt_tmp$'ms-gf:specevalue'),]
tmp_nona <- tmp_nona[order(tmp_nona$MSGFDB_SpecEValue, decreasing = F),]
bt_tmp <- bt_tmp[!is.na(bt_tmp$'pepseq'),]
```
There are ```r nrow(tmp)``` rows remained.

### Differences
Data from Docker is matched with from PNNL using scan number as ID. `r names(bt_tmp)[2:3]` are values from PNNL. `r names(bt_tmp)[4:5]` are the corresponding values from Docker Container. There are `r nrow(bt_tmp)` of rows in total with   `r nrow(bt_tmp[((bt_tmp$'ms-gf:specevalue'==bt_tmp$MSGFDB_SpecEValue) & (bt_tmp$pepseq==bt_tmp$Peptide)),])` exact match.

Aside from the `r nrow(tmp_nona)` unmatched values, `r sum(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue)` have different spec-evalue, `r sum(bt_tmp$pepseq!=bt_tmp$Peptide)` of different peptide identification.

```{r echo=F}
print("==Head==")
head(bt_tmp,2)
print("==Unmatched==")
head(tmp_nona,3)
print("==Tail==")
tail(bt_tmp,2)


print("==Unmatched==")
head(tmp_nona,2)
print("==Different Spec Values==")
head(bt_tmp[(bt_tmp$'ms-gf:specevalue'!=bt_tmp$MSGFDB_SpecEValue),],2)
print("==Different Peptides identification==")
head(bt_tmp[(bt_tmp$pepseq!=bt_tmp$Peptide),],2)

```